### 2019-04-03 README.Rmd
Moving table processing to data-raw
text changes
Completing country coverage in cites_parties()
README updates, begin vignette revision

**citesdb** is a package to conveniently analyze the full CITES shipment-level wildlife trade database, available at <https://trade.cites.org/>. This data consists of over 20 million records of reported shipments of wildlife and wildlife
products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).

@@

options(width=120)
knitr::opts_chunk$set(cache = TRUE)
citesdb::cites_disconnect()
MonetDBLite::monetdblite_shutdown()
```

@@

Once you fetch the data you can connect to the database with the `cites_db()`
command.  You can use the
`cites_shipments()` command to load a remote `tibble` that is backed by the
database but not loaded into R.  You can use this
to analyze CITES data without ever loading it into memory, then gather your
results with `collect()`. For example:

@@

cites_shipments() %>%
  group_by(Year) %>%
  summarize(number = n()) %>%
  arrange(desc(Year)) %>%
  collect()

@@

(_Note that running `collect()` on all of `cites_shipments()` will load a >3GB data frame into memory!_)

@@

If you don't need any of the bells and whistles of this package, you can download
the raw data as a single compressed TSV from the [releases page](https://github.com/ecohealthalliance/citesdb/releases),
or as a `.zip` file of many CSV files from original source at <https://trade.cites.org/>.

@@

The database also contains tables of field metadata, codes used, and CITES countries. This information
comes from ["A guide to using the CITES Trade Database"](https://trade.cites.org/cites_trade_guidelines/en-CITES_Trade_Database_Guide.pdf),
on the CITES website.  Convenience functions `cites_metadata()`, `cites_codes()`, and `cites_parties()`
access this information:

@@

More information on the release of shipment-level data can be found in the
`?guidance` help file.

### Basic analyses

```{r warning = FALSE}

library(ggplot2)
library(cowplot)

breaks <- seq(from = 1975, to = 2020, by = 5)

plot1 <- cites_shipments() %>%
  group_by(Year) %>%
  summarize(n_records = n()) %>%
  mutate(log10_n_records = log10(n_records)) %>%
  ggplot(aes(x = Year, y = n_records)) +
  geom_point() +
  geom_line(linetype = "dashed", size = 0.2) +
  ylab("Number of Records") +
  theme_minimal() +
  scale_x_continuous(breaks = breaks)

plot2 <- cites_shipments() %>%
  group_by(Year) %>%
  summarize(n_records = n()) %>%
  mutate(log10_n_records = log10(n_records)) %>%
  ggplot(aes(x = Year, y = log10_n_records)) +
  geom_point() +
  geom_line(linetype = "dashed", size = 0.2) +
  ylab("log10(Number of Records)") +
  ylim(1.5, 6.5) +
  theme_minimal() +
  scale_x_continuous(breaks = breaks)

plot_grid(plot1, plot2, nrow = 2)
```

@@

cites_disconnect()
MonetDBLite::monetdblite_shutdown()
```

@@

Want have feedback or want to contribute? Great! Please take a look at the [contributing guidelines](https://github.com/ecohealthalliance/citesdb/blob/master/.github/CONTRIBUTING.md) before filing an issue or pull request.

>>>>

**citesdb** is an R package to conveniently analyze the full CITES shipment-level wildlife trade database, available at <https://trade.cites.org/>. This data consists of over 20 million records of reported shipments of wildlife and wildlife products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).

@@

options(width = 120)
knitr::opts_chunk$set(cache = TRUE)
citesdb::cites_disconnect()
MonetDBLite::monetdblite_shutdown()
```

@@

Once you fetch the data you can connect to the database with the `cites_db()` command. You can use the `cites_shipments()` command to load a remote `tibble` that is backed by the database but not loaded into R. You can use this to analyze CITES data without ever loading it into memory, then gather your results with the `dplyr` function `collect()`. For example:

@@

cites_shipments() %>%
  group_by(Year) %>%
  summarize(n_records = n()) %>%
  arrange(desc(Year)) %>%
  collect()

@@

(_Note that running `collect()` on all of `cites_shipments()` will load a >3 GB data frame into memory!_)

@@

If you don't need any of the bells and whistles of this package, you can download the raw data as a single compressed TSV file from the [releases page](https://github.com/ecohealthalliance/citesdb/releases), or as a `.zip` file of many CSV files from original source at <https://trade.cites.org/>.

@@

The package database also contains tables of field metadata, codes used, and CITES countries. This information comes from ["A guide to using the CITES Trade Database"](https://trade.cites.org/cites_trade_guidelines/en-CITES_Trade_Database_Guide.pdf), on the CITES website. Convenience functions `cites_metadata()`, `cites_codes()`, and `cites_parties()` access this information:

@@

@@

citesdb::cites_disconnect()
MonetDBLite::monetdblite_shutdown()
```

More information on the release of shipment-level CITES data can be found in the `?guidance` help file.

@@

Have feedback or want to contribute? Great! Please take a look at the [contributing guidelines](https://github.com/ecohealthalliance/citesdb/blob/master/.github/CONTRIBUTING.md) before filing an issue or pull request.
